Investigating the Impact of Macroeconomic Indicators on the FTSE 250 Index Using Statistical and Machine Learning Methods

### 2024-08-26 tex/paper.tex
final revisions

\begin{abstract}
    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate, and the money supply (M3) measure on the price of the FTSE 250 Index (FTSE250).
    In addition, it assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables. This study finds quantitative estimates for the short-term and long-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index. It also demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 Index.
    \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, radial basis function neural network, time-series forecasting}
\end{abstract}

@@

Macroeconomic indicators can also be effectively used in machine learning 
methods to forecast stock prices by serving as input features that capture 
the economic environment’s influence on stock market behaviour. By carefully 
selecting the most relevant indicators and preprocessing data, 
as well as choosing appropriate models, machine learning can enhance the accuracy of stock price forecasts (\citeboth{Prasad2023}).

@@

In order to analyse the dynamic relationship between variables in time series data, this study
will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error
Correction Model (ECM). This research also employs a radial basis function neural network
(RBFNN) to predict the price movements of the FTSE250.

The remainder of this paper will be organised as follows. 
Section \ref{sec:lit} provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
Section \ref{sec:meth} presents the methodology of the paper. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
the study.

@@

\subsection{Data Sources}
Monthly FTSE 250 Index price data, 
monthly CPI data, monthly M3 money supply data, monthly USD/GBP exchange 
rate data and interest rate data between Oct 1992 to March 2024 has been 
utilised for this study giving a dataset with 379 observations. The sources 
are presented in Table \ref{table: tabl}.

\clearpage
\begin{table}[h!]
    \centering
    \caption{Data sources}
    \label{table: tabl}
    \begin{tabular}{lll}
        \toprule
        \textbf{Variable} & \textbf{Notation} & \textbf{Source} \\
        \midrule
        FTSE 250 Index & FTSE$_{250}$ & Yahoo Finance \\
        Consumer Price Index & CPI & Office for National Statistics (ONS) \\
        USD/GBP Exchange Rate (£) & EXCHG & Federal Reserve of St Louis \\
        Interest Rate ($\%$) & INT & Bank of England \\
        Money Supply (M3) (millions £) & M3 & Bank of England \\
        \bottomrule
    \end{tabular}
\end{table}

\subsection{Data Preprocessing}
For the interest rate data, it was necessary to address an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
does not convene on a monthly basis, and are not obligated to 
announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using Algorithm \ref{alg:interest_rate_adjustment}.

@@

Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
increasing role played by monetary policy in the UK since May 1997 when the Bank of England was given operational responsibility for setting interest rates (\citeboth{king1997changes}). The Bank of England was forced to adjust interest rates 
several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
post-COVID inflationary pressures.

@@

The ARDL model specifies the dependent variable as a function of its own 
lagged values and the lagged values of independent variables. 
The variables are permitted to be of differing stationarity. The stationarity 
of the variables was tested using the Augmented Dickey-Fuller 
Unit Root Test found within the tseries package in R. The optimal lags for each of the 
variables were calculated using the Schwarz Information Criterion, contained
within the auto\_ardl function in the ARDL package in R.

@@

The intercept term was given by $\alpha_0$. The 
remaining $\alpha_i, \beta_i$ coefficients represent the long-run relationships 
between the variables. The coefficients were computed using the ardl function in the 
ARDL package in R.

@@

Here, the intercept term was given by $\gamma_0$. The 
remaining $\gamma_i, \delta_i$ coefficients represent the short-run 
relationships between the variables. The $\lambda$ coefficient 
describes the speed at which short-run deviations from the long-run equilibrium
are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that guides the deviation 
to the long-run equilibrium. These coefficients, as well 
as the error correction term ECT, were computed using the recm function in the 
ARDL package in R.

Subsequently, cointegration was checked. The null and alternate hypothesis were:

@@

for all $i = 1,2,3,4$. The bounds\_f\_test in 
the ARDL library, based on the bounds test approach 
outlined in \citeboth{pesaran2001}, was used to calculate the 
F-statistic of the ARDL model and subsequently test for cointegration.

Finally, diagnostic tests are conducted to confirm that the model does not
display autocorrelation, heteroskedascity, non-normality and that the cumulative 
sum of residuals and squared residuals are both stable. This is to 
ensure our regression is not spurious.

@@

Radial basis function neural networks are a type of artificial 
neural network that use radial basis functions at each node in the 
activation layer of the network. They are able to identify complex, 
non-linear relationships between time-series variables and have been used
extensively in this aspect. The advantages of radial basis function neural networks over other 
neural networks in time-series forecasting lies in their 
ability to operate effectively on small sample sizes.

@@

Neural networks work best with data that is specified between a specific 
range (\citeboth{mendelsohn1993}). Also, to achieve stationarity
it was required to take the first difference all of the variables. Therefore, all of the 
variables in the dataset described in were normalised using a z-score normalisation
and differenced, as described in Algorithm \ref{alg:z_score_differencing}.

@@

The inputs into the network are 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced, normalised dataset $Z$

@@

where \(\Delta \text{CPI}_i\), \(\Delta \text{EXCHG}_i\), \(\Delta \text{INT}_i\), and \(\Delta \text{M3}_i\) 
represent differenced monthly values in the CPI, USD/GBP exchange rate, interest rate, and M3 money supply, respectively.

The output is a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced monthly values in the FTSE250.

@@

Based on this, the train-test process was conducted:
\begin{enumerate}
    \item The optimal hyperparameters in the covariance function \eqref{eq: gimq} 
    are computed using GridSearch Cross Validation. These hyperparameters are:
    \begin{enumerate}
        \item $\alpha$: a regularisation parameter representing a small quantity
        that is added to the diagonals of the covariance matrix to ensure it is 
        invertible. 
        \item $\beta$: the power in the covariance function, which controls the shape of the function.
        \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for. It moderates the risk of overfitting and underfitting.
        \item Length scale: a parameter that controls the influence that each kernel is to have. Analogous to 
        the width each radial basis function neuron possesses.
    \end{enumerate}
    \item Training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) are successively passed through the 
    covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
    all training observations, $K$.
    \item A $4$-dimensional input space $X$ takes in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
    calculates the posterior weights distribution. 
    Each possible weight vector $w$ represents a different function, explicitly 
    based on the radial basis function covariance function, that could fit the data. In the 
    hidden layer the expectation is then taken, which
    combines all weights and finds the mean, weighted by the 
    respective probabilities. This represents the best functional estimate of the price movement. Based on this estimate 
    the predicted output generated.
    \item The testing process was conducted similarly. The 
    posterior weight distribution that contains the information of the training phase becomes the new 
    prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
    were calculated, and then the predictive output was found (\citeboth{rasmussen2006}).

@@

A short-term and long-term forecast is conducted using this process. For the short-term forecast a test set of 5 months was used, and for the long-term forecast a test set of 57 months was used.

@@

\begin{table}[h!]
    \centering
    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test\tablefootnote{All of the variables are tested in natural logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
    at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}}
    \begin{tabular}{lccc}
        \toprule
        \textbf{Variable} & \textbf{Level ADF} & \textbf{First Difference ADF} & \textbf{Stationarity} \\
        \midrule
        FTSE$_{250}$ & $-3.2646^{****}$ & $-6.6892^{*}$ & At first difference \\
        CPI          & $-1.4594$ & $-4.705^{*}$ & At first difference \\
        EXCHG        & $-2.2467$ & $-7.3855^{*}$ & At first difference \\
        INT          & $-2.0946$ & $-4.1156^{*}$ & At first difference \\
        M3           & $-0.70827$ & $-6.3806^{*}$ & At first difference \\
        \bottomrule
    \end{tabular}
\end{table}

The optimum lags for the ARDL model were ($FTSE_{250} = 5$, $CPI = 4$, $EXCHG=7$, $INT =7$, $M3=4$) for  The resulting F-statistic of $4.1$ was compared against the critical values
given by \citeboth{pesaran2001}, providing evidence of cointegration 
at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
level.

@@

The interest rate has a significantly
negative short-term impact on the FTSE250 price in the short term. 
This can be attributed to the immediate effect of interest rate hikes, 
which decrease the present value of future earnings according to the 
Discounted Cash Flow (DCF) valuation model, thereby 
affecting investor sentiment and depressing stock prices. 
This finding is consistent with the studies by 
\citeboth{alam2009}, \citeboth{demir2019}, and \citeboth{neifar2023}.

The money supply has a significantly negative short-term effect on the 
FTSE250 price, as noted by \citeboth{olawale2014}. 
According to \citeboth{sellin2001}, this supports the Keynesian argument 
that an increase in money supply leads to expectations of a future 
monetary contraction. This expectation increases the demand for 
funds, driving up current interest rates and reducing stock prices 
through the same mechanism as described previously.

In the short-term, before structural shifts can materialise, the
weakening of the Pound against the Dollar has a positive, but 
insignificant effect. As the USD has been 
the global reserve currency for many decades, the depreciation of the Pound 
against it will lead to increased revenue in Pound terms for FTSE250 
companies with substantial overseas operations. 
This is because their foreign earnings, 
when converted back into Pounds, will be worth more, 
potentially increasing profits. However, the insignificant impact suggests 
that FTSE250 firms are resilient to short-term fluctuations in the stock market 
price. The possible reasons for this are that they have diversified 
operations, meaning fluctuations against the Dollar will not threaten 
their profits and they use hedging strategies involving financial instruments 
such as currency forwards which protect against volatility.

The insignificance of inflation, defined as a positive 
increase in the CPI, confirms the findings of 
\citeboth{gultekin1983}, \citeboth{firth1979} and \citeboth{neifar2023}. 
The CPI as a measure does not uniformly represent all sectors of the economy, 
and may consequently be unrepresentative of inflationary changes in the 
sectors that predominate the FTSE250. In addition, different sectors 
are more vulnerable to inflation than others which may distort the effect 
of inflation even further. Moreover, inflation may be anticipated 
by investors and already be factored into stock prices.

The error correction term 
was negative and statistically significant ensuring that long-run equilibrium 
can be achieved. The adjustment speed of $0.1794$ indicates that 
roughly $18\%$ of the fluctuation in the previous month's price caused by the 
macroeconomic variable movements is adjusted back to the long-run equilibrium in the subsequent month. This implies 
that it takes approximately 6 months for the FTSE250 to return to its long-run equilibrium following 
changes in the selected macroeconomic variables.

@@

The insignificant long-term effect of the money supply can be explained by several reasons.
Firstly, as discussed in \citeboth{sellin2001}, it is possible that the FTSE250 is efficient in processing changes in the 
money supply, at least over certain intervals if not always. This would mean anticipated changes in the money supply would not affect the stock market price, only the
unanticipated component of a change in money supply would, as established by \citeboth{sorensen1982} and \citeboth{bernanke2005}. 
It is also possible that the M3 measure is too broad of a measure and includes overly illiquid components that are not as impactful 
on the stock market dynamics.

@@

The short-term prediction yielded MAE = $0.07$ and 
$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figure \ref{fig:shortmonthly}.
The significantly better performance of the model in the short-run
indicates that the selected macroeconomic variables possess
a stronger predictive power over shorter periods. This was highlighted 
earlier in the ARDL model, with the interest rate and money supply significantly
impacting the FTSE250 price in the short-run.

@@

Stock markets provide important information on the economy, 
and economic policymakers are vigilant in tracking the fluctuations 
in these markets, and try to forecast the 
forthcoming fluctuations based on policy changes that affect the macroeconomy.
Within this context, the study examined the 
impacts of the interest rate, USD/GBP
exchange rate, the Consumer Price Index, and 
M3 money supply market activity on
the FTSE250 over the October 1993 – April 2024 period.
The results suggest a cointegrated long-run
relationship between the macroeconomic variables and the stock market index. 
The long-run coefficients imply that
increases in the interest rate and depreciation in the exchange rate
significantly depress the FTSE250 value. 
Policymakers dealing with forecasting the UK stock markets
should consider the factors in the study, and to raise the stock
market's value, a lower interest rate regime should be adopted. This will 
stimulate consumer activity, reduce operating costs for firms, and make 
stocks more attractive investments, thereby increasing the stock market value. 
In addition, the Pound should be increased in value to lower import costs 
and increase the real value of firm earnings, both of which should drive up 
the value of the stock market. The combined effect of these changes will 
create a more stimulating business environment by increasing the capital 
available to firms for research and development, and more efficient 
capital and labour, in turn 
driving economic growth and prosperity. However, 
as with all economic policy shifts, these changes may have 
unintended consequences on other areas of the economy, potentially 
undermining any positive changes. Any policy changes must cautiously be made 
with this at the forefront.

>>>>

\begin{abstract}
    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate, and the money supply (M3) measure on the price of the FTSE 250 Index (FTSE250).
    In addition, it assesses the effectiveness of radial basis function neural networks (RBFNN) as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables. This study finds quantitative estimates for the short-term and long-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index. It also demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 Index.
    \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, radial basis function neural network, time-series forecasting}
\end{abstract}

@@

Macroeconomic indicators can also be effectively used in machine learning 
methods to forecast stock prices by serving as input features that capture 
the economic environment’s influence on stock market behaviour. By carefully 
selecting the most relevant indicators, 
as well as choosing the appropriate models, machine learning can enhance the accuracy of stock price forecasts (\citeboth{Prasad2023}).

@@

In order to analyse the dynamic relationship between the variables in the time series data, this study
will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error
Correction Model (ECM). This research also employs a radial basis function neural network to predict the price movements of the FTSE250.

The remainder of this paper will be organised as follows. 
Section \ref{sec:lit} provides a brief review of the literature in this problem space and explains the rationale behind the selection of macroeconomic variables.
Section \ref{sec:meth} presents the methodology. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
the study.

@@

\subsection{Data Sources}

The study uses monthly FTSE250 price data between October 1992 and March 2024,
thus providing 379 observations. The data for the macroeconomic indicators 
was also captured on a monthly basis in the same time frame, except for the 
interest rate. This is because the Bank of England's Monetary Policy 
Committee, 
which is responsible for setting interest rates, meets on an \textit{ad hoc} 
basis\footnote{The MPC typically meets roughly every six weeks, however, in uncertain economic 
climes it may convene far more frequently.}
to decide on whether the interest rate should be changed. The sources 
are presented in Table \ref{table: tabl}.

\clearpage
\begin{table}[h!]
    \centering
    \caption{Data sources}
    \label{table: tabl}
    \begin{tabular}{llll}
        \toprule
        \textbf{Variable} & \textbf{Frequency} & \textbf{Notation} & \textbf{Source} \\
        \midrule
        FTSE 250 Index & Monthly & FTSE$_{250}$ & Yahoo Finance \\
        Consumer Price Index & Monthly & CPI & Office for National Statistics (ONS) \\
        USD/GBP Exchange Rate (£) & Monthly & EXCHG & Federal Reserve of St Louis \\
        Interest Rate ($\%$) & Variable & INT & Bank of England \\
        Money Supply (M3) (millions £) & Monthly & M3 & Bank of England \\
        \bottomrule
    \end{tabular}
\end{table}

\subsection{Data Preprocessing}
\label{section: prep}

To address the variability in the interest data, Algorithm \ref{alg:interest_rate_adjustment}
was used to convert the data into a monthly format, ensuring a uniform
frequency across the variables.

@@

Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
increasing role played by monetary policy in the UK since May 1997 when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}). The Bank of England was forced to adjust interest rates 
several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
post-COVID inflationary pressures.

@@

The ARDL model specifies the dependent variable as a function of its own 
lagged values and the lagged values of independent variables. 
The variables are permitted to be of differing stationarity. The stationarity 
of the variables was tested using the Augmented Dickey-Fuller 
Unit Root Test found within the \texttt{tseries} package in R. The optimal lags for each of the 
variables were calculated using the Schwarz Information Criterion, contained
within the auto\_ardl function in the ARDL package in R.

@@

The intercept term was given by $\alpha_0$. The 
remaining $\alpha_i, \beta_i$ coefficients represented the long-run relationships 
between the variables. The coefficients were computed using the \texttt{ardl} function in the 
\texttt{ARDL} library in R.

@@

The intercept term was given by $\gamma_0$. The 
remaining $\gamma_i, \delta_i$ coefficients represented the short-run 
relationships between the variables. The $\lambda$ coefficient 
described the speed at which short-run deviations from the long-run equilibrium
were adjusted. A negative $\gamma$ implies that there existed a correction mechanism that guided the deviation 
back to the long-run equilibrium. These coefficients, as well 
as the error correction term ECT, were computed using the \texttt{recm} function in the 
\texttt{ARDL} library in R.

The \texttt{{bounds\_f\_test}} in 
the \texttt{ARDL} library, based on the bounds test approach 
outlined in \citeboth{pesaran2001}, was used to calculate the 
F-statistic of the ARDL model and subsequently test for cointegration.

@@

for all $i = 1,2,3,4$.

To confirm that the model did not display autocorrelation, heteroskedasticity, non-normality, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted, respectively

@@

This study employed radial basis function neural networks, 
a type of artificial neural network that utilise radial basis functions 
at each node in the activation layer. 
This was because radial basis function neural networks 
are well-suited for identifying complex, non-linear relationships 
between time-series variables, making them particularly valuable in 
forecasting tasks. In addition, radial basis function neural networks
have been shown to be effective in operating on small sample sizes, 
which was critical for the data constraints of this study.

@@

Neural networks work best with data that is specified between a specific 
range (\citeboth{mendelsohn1993}).
Also, based on the unit root tests in Table \ref{table: unitroot} the variables in the dataset $X$, described in \ref{section: prep},
only achieve stationary at first difference.
Therefore, all of the 
variables in the dataset $X$ were normalised using a z-score normalisation and differenced,
as described in Algorithm \ref{alg:z_score_differencing}.

@@

The inputs into the network were 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced, normalised dataset $Z$

@@

where \(\Delta \text{CPI}_i\), \(\Delta \text{EXCHG}_i\), \(\Delta \text{INT}_i\), and \(\Delta \text{M3}_i\) 
represented differenced monthly values in the CPI, USD/GBP exchange rate, interest rate, and M3 money supply, respectively.

The output was a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced monthly values in the FTSE250.

@@

The approach outlined in the \hyperref[sec: appendix]{Appendix} was used to conduct the train-test process:
\begin{enumerate}
    \item The optimal hyperparameters in the covariance function \eqref{eq: gimq} 
    were computed using GridSearch Cross Validation. These hyperparameters were:
    \begin{enumerate}
        \item $\alpha$: a regularisation parameter representing a small quantity
        that was added to the diagonals of the covariance matrix to ensure it was 
        invertible. 
        \item $\beta$: the power in the covariance function, which controlled the shape of the function.
        \item $\sigma$: a parameter which controlled how much of the noise in the data the kernel accounted for. It moderated the risk of overfitting and underfitting.
        \item Length scale: a parameter that controlled the influence that each kernel was to have. Analogous to 
        the width that each radial basis function neuron possesses.
    \end{enumerate}
    \item Training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) were successively passed through the 
    covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
    all training observations, $K$.
    \item A $4$-dimensional input space $X$ took in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
    calculated the posterior weights distribution. 
    Each possible weight vector $w$ represented a different function, explicitly 
    based on the radial basis function covariance function, that could fit the data. In the 
    hidden layer the expectation of this distribution was then taken, which
    combined all the weights and found the mean, weighted by the 
    respective probabilities. This represented the best functional estimate of the price movement. Based on this estimate 
    the predicted output was generated.
    \item The testing process was conducted similarly. The 
    posterior weight distribution that contained the information of the training phase became the new 
    prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
    were calculated, and then the predictive output was found (\citeboth{rasmussen2006}).

@@

A short-term and long-term forecast was conducted using this process. For the short-term forecast a test set of 5 months was used, and for the long-term forecast a test set of 57 months was used.

@@

\begin{table}[h!]
    \centering
    \label{table: unitroot}
    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test\tablefootnote{All of the variables are tested in natural logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
    at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}}
    \begin{tabular}{lccc}
        \toprule
        \textbf{Variable} & \textbf{Level ADF} & \textbf{First Difference ADF} & \textbf{Stationarity} \\
        \midrule
        FTSE$_{250}$ & $-3.2646^{****}$ & $-6.6892^{*}$ & At first difference \\
        CPI          & $-1.4594$ & $-4.705^{*}$ & At first difference \\
        EXCHG        & $-2.2467$ & $-7.3855^{*}$ & At first difference \\
        INT          & $-2.0946$ & $-4.1156^{*}$ & At first difference \\
        M3           & $-0.70827$ & $-6.3806^{*}$ & At first difference \\
        \bottomrule
    \end{tabular}
\end{table}

The optimum lags for the ARDL model were ($FTSE_{250} = 5$, $CPI = 4$, $EXCHG=7$, $INT =7$, $M3=4$). The resulting F-statistic of $4.1$ was compared against the critical values
given by \citeboth{pesaran2001}, providing evidence of cointegration 
at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
level.

@@

The interest rate had a significantly
negative short-term impact on the FTSE250 price in the short term. 
This can be attributed to the immediate effect of interest rate hikes, 
which decrease the present value of future earnings according to the 
Discounted Cash Flow (DCF) valuation model, thereby 
affecting investor sentiment and depressing stock prices. 
This finding was consistent with the studies by 
\citeboth{alam2009}, \citeboth{demir2019}, and \citeboth{neifar2023}.

The money supply had a significantly negative short-term effect on the 
FTSE250 price, as noted by \citeboth{olawale2014}. 
According to \citeboth{sellin2001}, this aligned with the Keynesian argument 
that an increase in money supply leads to expectations of a future 
monetary contraction. This expectation increases the demand for 
funds, driving up current interest rates and reducing stock prices 
through the same mechanism as described previously.

In the short-term, before structural shifts can materialise, the
weakening of the Pound against the Dollar had a positive, but 
insignificant effect. As the USD has been 
the global reserve currency for many decades, the depreciation of the Pound 
against it would lead to increased revenue in Pound terms for FTSE250 
companies with substantial overseas operations. 
This is because their foreign earnings, 
when converted back into Pounds, will be worth more, 
potentially increasing profits. However, the insignificant impact suggests 
that FTSE250 firms are resilient to short-term fluctuations in the stock market 
price. The possible reasons for this are that they have diversified 
operations, meaning fluctuations against the Dollar will not threaten 
their profits and they use hedging strategies involving financial instruments 
such as currency forwards which protect against volatility.

The insignificance of inflation, defined as a positive 
increase in the CPI, confirmed the findings of 
\citeboth{gultekin1983}, \citeboth{firth1979} and \citeboth{neifar2023}. 
The CPI as a measure does not uniformly represent all sectors of the economy, 
and may consequently be unrepresentative of inflationary changes in the 
sectors that predominate the FTSE250. In addition, different sectors 
are more vulnerable to inflation than others which may distort the effect 
of inflation even further. Moreover, inflation may be anticipated 
by investors and already be factored into stock prices.

The error correction term 
was negative and statistically significant ensuring that long-run equilibrium 
can be achieved. The adjustment speed of $0.1794$ indicated that 
roughly $18\%$ of the fluctuation in the previous month's price caused by the 
macroeconomic variable movements was adjusted back to the long-run equilibrium in the subsequent month. This implies 
that it takes approximately 6 months for the FTSE250 to return to its long-run equilibrium following 
changes in the selected macroeconomic variables.

@@

The insignificant long-term effect of the money supply can be explained by several factors.
Firstly, as discussed in \citeboth{sellin2001}, it is possible that the FTSE250 is efficient in processing changes in the 
money supply, at least over certain intervals if not always. This would mean anticipated changes in the money supply would not affect the stock market price, only the
unanticipated component of a change in money supply would, as established by \citeboth{sorensen1982} and \citeboth{bernanke2005}. 
It is also possible that the M3 measure is too broad of a measure and includes overly illiquid components that are not as impactful 
on the stock market dynamics.

@@

The short-term prediction yielded MAE = $0.07$ and 
$R^2 = 61\%$ for a 5-month time horizon. The results obtained are presented in Figure \ref{fig:shortmonthly}.
The significantly better performance of the model in the short-run
indicates that the selected macroeconomic variables possess
a stronger predictive power over shorter periods. This was highlighted 
earlier in the ARDL model, with the interest rate and money supply significantly
impacting the FTSE250 price in the short-run.

@@

Stock markets provide important information on the economy, 
and economic policymakers are vigilant in tracking the fluctuations 
in these markets, and try to forecast the 
forthcoming fluctuations based on policy changes that affect the macroeconomy.
Within this context, the study examined the 
impacts of the interest rate, USD/GBP
exchange rate, the Consumer Price Index, and 
M3 money supply market activity on
the FTSE250 over the October 1993 – April 2024 period.
The results suggested a cointegrated long-run
relationship between the macroeconomic variables and the stock market index. 
The long-run coefficients implied that
increases in the interest rate and depreciation in the exchange rate
significantly depress the FTSE250 value. 
Policymakers dealing with forecasting the UK stock markets
should consider the factors in the study, and to raise the stock
market's value, a lower interest rate regime should be adopted. This will 
stimulate consumer activity, reduce operating costs for firms, and make 
stocks more attractive investments, thereby increasing the stock market value. 
In addition, the Pound should be increased in value to lower import costs 
and increase the real value of firm earnings, both of which should drive up 
the value of the stock market. The combined effect of these changes will 
create a more stimulating business environment by increasing the capital 
available to firms for research and development, and more efficient 
capital and labour, in turn 
driving economic growth and prosperity. However, 
as with all economic policy shifts, these changes may have 
unintended consequences on other areas of the economy, potentially 
undermining any positive changes. Any policy changes must cautiously be made 
with this at the forefront.
